Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-02-11 doc/invt.tex
invt.tex: add section for stabilization

@@

\begin{abstract}
  The Wang-Landau (WL) algorithm and metadynamics
  are two closely-related techniques in free energy simulations.
  %
  They allow one to sample a flat distribution
  along a quantity of interest, $z$,
  by adaptively constructing a bias potential
  that offsets the potential of mean force (PMF).
  %
  The two algorithms differ by
  the manner of updating the bias potential:
  %
  in the WL case, each simulation step triggers
  an update of the bias potential to only
  the bin containing the current $z$,
  while in the metadynamics case,
  the bias potential of several neighboring bins
  are affected as well.
  %
  The asymptotic error of the resulting PMF
  is determined by the updating magnitude,
  and the rate of the reduction of the magnitude
  over the simulation time.
  %
  In the WL case,
  a simple and effective formula for the magnitude
  is given by the inverse of the number of simulation steps.
  %
  In this study, we shall reexamine the optimality of this prescription,
  and attempt to extend it to the metadynamics case.
  %
  Further, we compared the two algorithms, and found that
  although metadynamics produces a smoother PMF profile,
  it is less effective in reducing errors with the shorter wavelengths.
  %
  Thus, the latter relies more heavily on the assumption of
  the smoothness of the PMF profile,
  especially with a wide updating window.
\end{abstract}

@@

A better solution is to artificially alter
the target distribution,
such that we can, in a single simulation, sample
a flat distribution\cite{mezei1987, berg1992, lee1993,
wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}
along $z$ over a wide range.
%
This alleviates the above problem of local trapping
as the system is now forced to travel more flexibly
along the $z$ direction,
which often represents a slow reaction coordinate.
%
It, however, requires one to efficiently construct a
bias potential for an often unknown system.
%
Note that the bias potential must exactly cancel the PMF
to achieve a flat distribution.

The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
and metadynamics\cite{huber1994, laio2002, laio2008, barducci2011, sutto2012}
are two widely-used and closely-related\cite{micheletti2004}
techniques for this purpose,
in which a bias potential is actively built-up on-the-fly
during simulation.
%
In essence, these techniques
encourage the system to leave a region that has been visited
by elevating the bias potential there.
%
A main difference of the two techniques
is that in the WL case, an update is limited to the bin
containing the current $z$,
while metadynamics adopts an extended window
covering several neighboring bins.
%
The latter ensures a smooth profile
suitable for numerical differentiation for force calculation,
and hence is more often used molecular dynamics simulations.
%
To be effective, the updates to bias potential
are made at every few steps,
while a side effect is that the underlying dynamics
is no longer a equilibrium sampling
required by statistical mechanics.
%
Thus, one has to reduce the magnitude of updating
over the course of simulation.

Naturally, the manner of reducing
the updating magnitude\cite{liang2007,
belardinelli2007, belardinelli2007jcp, belardinelli2008,
morozov2007, zhou2008, morozov2009,
komura2012, caparica2012, caparica2014,
barducci2008, dickson2011, dama2014}
determines the precision of the final bias potential,
hence that of PMF.
%
For the WL algorithm, it is found
that asymptotically the optimal magnitude of updating $\ln f$
should be given by the
inverse time\cite{liang2007,
belardinelli2007, belardinelli2007jcp, belardinelli2008,
morozov2007, zhou2008},
where ``time'' is defined as
the number of simulation steps
divided by the number of bins.

In this study,
we shall give another proof of the inverse-time formula.
%
We then try to derive a similar result
for the optimal updating magnitude
for updating schemes involving several neighboring bins,
such as those used in metadynamics.
%
It turns out that the optimal updating magnitude
depends nontrivially on the simulation length
and other parameters in the latter case.
%
We also shall show that
the single-bin scheme used by the WL algorithm
is best in asymptotic convergence,
although practical simulations may not need
to reach this regime.
%
The article is organized as follows.
%
We present the analytical results in Sec. \ref{sec:theory},
numerically verify some key aspects
in Sec. \ref{sec:results},
and conclude the article
in Sec. \ref{sec:conclusion}.

@@

We now consider a class of commonly-used updating scheme
for sampling a uniform distribution, $p_i = 1/n$.
%
These schemes are homogeneous,
or translational invariant with respect to $z$.
%
This invariance is natural for a periodic variable\cite{dama2014}.
%
Below, however, we focus on a variable
under the reflective boundary conditions.

@@

This matrix, being symmetric to the indices $i$ and $j$,
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution $p_i = 1/n$.
%
It has $n$ orthonormal eigenvectors,
$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
given by
%
\begin{equation}
\varphi^{(k)}_i
= \sqrt{
    \frac{ 2 - \delta_{k, 1} }
         {       n           }
       }
  \cos \frac{ \left( i - \frac 1 2 \right) (k - 1) \, \pi}
            {                  n                         }.
\label{eq:wband_eigenvector}
\end{equation}
%
with eigenvalues
%
\begin{align}
  \lambda_k
  &=
  \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}
  \label{eq:wband_eigenvalue1}
  \\
  &=
  1 - \sum_{l = 1}^b 4 \, \mu_l \sin^2 \frac{(k - 1)  \, l \pi}{ 2 \, n }
  .
  \label{eq:wband_eigenvalue}
\end{align}

@@

We can define a discrete kernel function from the set of numbers as
$$
\hat \mu_i = \mu_{ |i| }.
$$
Then, according to Eq. \eqref{eq:vt_diffeq_mbin},
the rate of change of the bias potential, $\dot v_i(t)$,
is given by the convolution of the above kernel, $\hat \mu_j$,
and the normalized histogram, $h_j/p_j$\cite{bussi2006},
and Eq. \eqref{eq:wband_eigenvalue1}
shows that the eigenvalues, $\lambda_k$, of the updating scheme
is the Fourier transform of the kernel.
%
So, for the updating scheme to be stable,
the Fourier transform of the kernel
must be nonnegative.

@@

For the one-step sampling,
we find from Eq. \eqref{eq:Gamma_onestep}
that $\Gamma(p) = \cot^2 p$
is peaked around $p \approx 0$,
%
Then
$$
\begin{aligned}
\Err_A
&
\propto
\frac{   2 \, n }
     { \pi \, t }
\left.
\frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
     { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
\right|_{ p \approx 0 }
\notag \\
&
\approx
\frac{   2 \, n }
     { \pi \, t }
\frac{             1             }
     {  \lambda \, (2 - \lambda) }
.
\end{aligned}
$$
%
The optimal $\lambda$ is therefore
%
\begin{equation}
\lambda \approx 1.
\label{eq:lambda_nn_onestep}
\end{equation}
%

\note{We can be a little more precise here.
%
From Eq. \eqref{eq:Gamma_onestep}, we get
%
$$
\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2),
$$
%
where we have introduced
$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
as a small parameter to avoid the divergence
around the origin.
%
$$
\begin{aligned}
\Err_A
&=
\frac{   2 \, n }
     { \pi \, t }
\int_0^{\pi/2}
    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
    \frac{ \cos^2 p }
         { \sin^2 p + \delta^2 }
\, dp
\notag \\
&
\stackrel{    \delta \ll 1     }
         { =\joinrel=\joinrel= }
\frac{   2 \, n }
     { \pi \, t }
\frac{             1             }
     {  \lambda \, (2 - \lambda) }
\frac{    1   }
     { \delta }
.
\end{aligned}
$$
The integral is evaluated as follows.
%
We again change variable $p \to \frac{ \pi } { 2 } - p$,
and
$$
\begin{aligned}
\Err_A
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\cdot
\frac{ \sin^2 p            }
     { \cos^2 p + \delta^2 }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\cdot
\frac{ 4 - \left( z + \frac 1 z \right)^2 }
     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
\, \frac{dz}{z}
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
\cdot
\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
\, \frac{ dz }{ z^3 }
\\
&=
\frac{ n } { \lambda \, t }
\sum_l \operatorname{Res}_l
,
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$,
and the integral is reduced to a sum of residues
around the poles within.

@@

\section{\label{eq:stabilize_wband}
Stabilization}

>>>>

\hypersetup{
  colorlinks,
  linkcolor={red!30!black},
  citecolor={green!20!black},
  urlcolor={blue!80!black}
}

@@

\begin{abstract}
  The Wang-Landau (WL) algorithm and metadynamics
  are two closely-related techniques in free energy simulations.
  %
  They allow one to sample a flat distribution
  along a quantity of interest, $z$,
  by adaptively constructing a bias potential
  that offsets the potential of mean force (PMF).
  %
  The two algorithms differ by
  the manner of updating the bias potential:
  %
  in the WL case, each simulation step triggers
  an update to only
  the bin containing the current $z$,
  while in the metadynamics case,
  several neighboring bins
  are affected as well.
  %
  The asymptotic error of the resulting PMF
  is determined by the updating schedule,
  or the updating magnitude
  as a function of simulation time.
  %
  In the WL case,
  a simple and effective formula for the magnitude
  is given by the inverse of the number of simulation steps.
  %
  In this study, we shall reexamine the optimality of this prescription,
  and attempt to give similar prescriptions for the metadynamics case.
  %
  Our results show that the optimal schedule
  and the effectiveness of metadynamics
  are sensitive to the simulation length and other parameters.
  %Further, we compared the two algorithms, and found that
  %although metadynamics produces a smoother PMF profile,
  %it is less effective in reducing errors with the shorter wavelengths.
  %%
  %Thus, the latter relies more heavily on the assumption of
  %the smoothness of the PMF profile,
  %especially with a wide updating window.
\end{abstract}

@@

A better solution is to artificially alter
the target distribution,
such that we can, in a single simulation, sample
a flat distribution\cite{mezei1987, berg1992, lee1993,
wang2001, wang2001pre, laio2002, laio2008, barducci2011, sutto2012}
along $z$ over a wide range.
%
This alleviates the above problem of local trapping
as the system is now forced to travel more flexibly
along the $z$ direction,
which often represents
a direction of slow motion.
%
It, however, requires one to efficiently construct a
bias potential for an often unknown system.
%
Note that the bias potential must exactly cancel the PMF
to achieve a flat distribution.

The Wang-Landau (WL) algorithm\cite{wang2001, wang2001pre}
and metadynamics\cite{huber1994, laio2002, laio2008, barducci2011, sutto2012}
are two widely-used and closely-related\cite{micheletti2004}
techniques for this purpose,
in which a bias potential is actively built-up on-the-fly
during simulation.
%
In essence, these techniques
encourage the system to leave a region that has been visited
by elevating the bias potential there.
%
A main difference of the two techniques
is that in the WL case, an update is limited to the bin
containing the current $z$,
while metadynamics adopts an extended updating window
covering several neighboring bins.
%
The latter ensures a smooth profile
suitable for numerical differentiation for force calculation,
and hence is more often used molecular dynamics simulations.
%
To be effective, the updates to bias potential
are made at every few steps,
while a side effect is that the underlying dynamics
is no longer a equilibrium sampling
required by statistical mechanics.
%
Thus, one often has to reduce the magnitude of updating
over the course of simulation.

Naturally, the updating schedule,
or the manner of reducing
the updating magnitude\cite{liang2007,
belardinelli2007, belardinelli2007jcp, belardinelli2008,
morozov2007, zhou2008, morozov2009,
komura2012, caparica2012, caparica2014,
barducci2008, dickson2011, dama2014}
determines the precision of the final bias potential,
hence that of PMF.
%
For the WL algorithm, it is found
that asymptotically the optimal magnitude of updating $\ln f$
should be given by the
inverse time\cite{liang2007,
belardinelli2007, belardinelli2007jcp, belardinelli2008,
morozov2007, zhou2008},
where ``time'' is defined as
the number of simulation steps
divided by the number of bins.

In this study,
we shall reexamine the inverse-time formula,
and similarly optimize schedules
for updating schemes involving several neighboring bins,
such as those used in metadynamics.
%
It turns out that the optimal updating schedule
can be sensitive the simulation length.
%
We also shall show that
the single-bin scheme used by the WL algorithm
is best in asymptotic convergence.
%
For finite-length simulations, however,
metadynamics may deliver better performance.
%
The article is organized as follows.
%
We present the analytical results in Sec. \ref{sec:theory},
numerically verify some key aspects
in Sec. \ref{sec:results},
and conclude the article
in Sec. \ref{sec:conclusion}.

@@

We now consider a class of commonly-used updating scheme
for sampling a uniform distribution, $p_i = 1/n$.
%
These schemes are homogeneous,
or translational invariant with respect to $z$.
%
This invariance is natural for a periodic variable\cite{dama2014}.
%
Below, however, we focus on a $z$ that is
under the reflective boundary conditions.

@@

This matrix, being symmetric to the indices $i$ and $j$,
satisfies Eqs. \eqref{eq:w_sumj}-\eqref{eq:w_balance},
for the flat distribution $p_i = 1/n$.
%
It has $n$ orthonormal eigenvectors,
$\pmb\varphi^{(1)}, \dots, \pmb\varphi^{(n)}$,
given by
%
\begin{equation}
\varphi^{(k)}_i
= \sqrt{
    \frac{ 2 - \delta_{k, 1} }
         {       n           }
       }
  \cos \frac{ \left( i - \frac 1 2 \right) (k - 1) \, \pi}
            {                  n                         }.
\label{eq:wband_eigenvector}
\end{equation}
%
with eigenvalues
%
\begin{align}
  \lambda_k
  &=
  \mu_0 + \sum_{l = 1}^b 2 \, \mu_l \cos \frac{(k - 1)  \, l \pi}{n}
  \label{eq:wband_eigenvalue}
  %\\
  %&=
  %1 - \sum_{l = 1}^b 4 \, \mu_l \sin^2 \frac{(k - 1)  \, l \pi}{ 2 \, n }
  %.
  %\label{eq:wband_eigenvalue2}
\end{align}

@@

We can define a discrete kernel function from the set of numbers as
$$
\hat \mu_i = \mu_{ |i| }.
$$
Then, according to Eq. \eqref{eq:vt_diffeq_mbin},
the rate of change of the bias potential, $\dot v_i(t)$,
is given by the convolution of the above kernel, $\hat \mu_j$,
and the normalized histogram, $h_j/p_j$\cite{bussi2006},
and Eq. \eqref{eq:wband_eigenvalue}
shows that the eigenvalues, $\lambda_k$, of the updating scheme
is the Fourier transform of the kernel.
%
So, for the updating scheme to be stable,
the Fourier transform of the kernel
must be nonnegative.
%
In Sec. \ref{sec:stabilize_wband},
we show how to stabilize an arbitrary kernel.

@@

For the one-step sampling,
we find from Eq. \eqref{eq:Gamma_onestep}
that $\Gamma(p) = \cot^2 p$
is peaked around $p \approx 0$,
%
Then
$$
\begin{aligned}
\Err_A
&
\propto
\frac{   2 \, n }
     { \pi \, t }
\left.
\frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
     { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
\right|_{ p \approx 0 }
\notag \\
&
\approx
\frac{   2 \, n }
     { \pi \, t }
\frac{             1             }
     {  \lambda \, (2 - \lambda) }
.
\end{aligned}
$$
%
The optimal $\lambda$ is therefore
%
\begin{equation}
\lambda \approx 1.
\label{eq:lambda_nn_onestep}
\end{equation}
%
\note{We can be a bit more precise here.
%
From Eq. \eqref{eq:Gamma_onestep}, we get
%
$$
\Gamma(p) = \cot^2 p \approx \cos^2 p / (\sin^2 p + \delta^2),
$$
%
where we have introduced
$\delta \propto \sin[ \pi / (2 \, n) ] \sim 1/n$
as a small parameter to avoid the divergence
around the origin.
%
$$
\begin{aligned}
\Err_A
&=
\frac{   2 \, n }
     { \pi \, t }
\int_0^{\pi/2}
    \frac{            \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2         }
         { \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
    \frac{ \cos^2 p }
         { \sin^2 p + \delta^2 }
\, dp
\notag \\
&
\stackrel{    \delta \ll 1     }
         { =\joinrel=\joinrel= }
\frac{   2 \, n }
     { \pi \, t }
\frac{             1             }
     {  \lambda \, (2 - \lambda) }
\frac{    1   }
     { \delta }
.
\end{aligned}
$$
The integral is evaluated as follows.
%
We again change variable $p \to \frac{ \pi } { 2 } - p$,
and
$$
\begin{aligned}
\Err_A
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\int_0^{2 \, \pi}
\frac{ \left(1 - 4 \, \mu_1 \, \cos^2 p \right)^2 }
     {       2 - 8 \, \mu_1 \, \cos^2 p - \lambda }
\cdot
\frac{ \sin^2 p            }
     { \cos^2 p + \delta^2 }
\, dp
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[1 -      \mu_1 \, \left(z + \frac{1}{z}\right)^2 \right]^2 }
     {       2 - 2 \, \mu_1 \, \left(z + \frac{1}{z}\right)^2 - \lambda }
\cdot
\frac{ 4 - \left( z + \frac 1 z \right)^2 }
     { \left( z + \frac 1 z \right)^2 + 4 \, \delta^2 }
\, \frac{dz}{z}
\\
&=
\frac{n}{\lambda \, t}
\frac{1}{2 \pi i}
\oint
\frac{ \left[z^2 -      \mu_1 \, (z^2 + 1)^2 \right]^2   }
     {  (2 - \lambda) \, z^2 - 2 \, \mu_1 \, (z^2 + 1)^2 }
\cdot
\frac{ 4 \, z^2 - ( z^2 + 1 )^2 }
     { ( z^2 + 1 )^2 + 4 \, \delta^2 \, z^2 }
\, \frac{ dz }{ z^3 }
\\
&=
\frac{ n } { \lambda \, t }
\sum_l \operatorname{Res}_l
,
\end{aligned}
$$
where the contour is the unit circle
in the complex plane of $z$,
and the integral is reduced to a sum of residues
around the poles within.

@@

\subsubsection{\label{sec:stabilize_wband}
Stabilization}

The updating scheme determined by an arbitrary set of numbers,
$\mu_1, \dots, \mu_b$, is not necessarily stable,
i.e., the eigenvalues from Eq. \eqref{eq:wband_eigenvalue}
can be negative.
%
Below we show how to minimally modify these numbers
to make the eigenvalues nonnegative.

The key of this modification is the inversion formula
%
\begin{align}
  \mu_l
  =
  \frac 1 n
  \left[
    \frac{ \lambda_1 + (-1)^l \, \lambda_{n+1} }
         {               2                     }
    +
    \sum_{ k = 1 }^{ n - 1 }
      \lambda_{ k + 1 } \,
      \cos \frac{ k \, l \, \pi }
                {      n        }
  \right],
  \label{eq:mu_from_lambda}
\end{align}
%
where
%
\begin{align}
  \lambda_{ n + 1 }
  =
  (-1)^{ n - 1 }
  \left[
    \lambda_1
    +
    2 \, \sum_{ k = 1 }^{ n - 1 }
            (-1)^k \, \lambda_{k + 1}
  \right],
  \label{eq:lambdan}
\end{align}
%
which ensures that the $\mu_n$
computed from Eq. \eqref{eq:mu_from_lambda}
vanishes.
%
\note{The inversion formula is derived from the cosine transform.
  %
  If we define $\mu_n = 0$ and
  %
  \begin{align}
    \mu_{ 2 n - l } = \mu_l
    \quad
    \mathrm{for\;} l = 1, \dots, n - 1,
    \label{eq:mu_reflection}
  \end{align}
  %
  then Eq. \eqref{eq:wband_eigenvalue} can be rewritten as
  %
  \begin{align}
    \lambda_{k+1}
    =
    \sum_{ l = 0 }^{ 2 \, n - 1 }
    \mu_l \, \cos \frac{ k \, l \, \pi } { n }.
    \label{eq:lambda_cosine_sum}
  \end{align}
  %
  This formula for $\lambda_{k+1}$
  can be readily extended to $k = 2 \, n - 1$,
  and we have
  %
  \begin{align}
    \lambda_{ 2 \, n + 1 - k } = \lambda_k.
    \label{eq:lambda_reflection}
  \end{align}
  %
  Thus,
  %
  $$
  \begin{aligned}
    \sum_{ k = 0 }^{ 2 \, n - 1 }
      \lambda_{ k + 1 } \,
      \cos \frac{ k \, p \, \pi }
                {      n        }
    &=
    \sum_{ k = 0 }^{ 2 \, n - 1 }
      \sum_{ l = 0 }^{ 2 \, n - 1 }
        \mu_l \,
        \cos \frac{ k \, l \, \pi }
                  {      n        }
        \cos \frac{ k \, p \, \pi }
                  {      n        }
    \\
    &=
    \sum_{ l = 0 }^{ 2 \, n - 1 }
      \frac{ \mu_l } { 2 }
      \sum_{ k = 0 }^{ 2 \, n - 1 }
        \cos \frac{ k \, (p + l) \, \pi }
                  {      n        }
                  +
        \cos \frac{ k \, (p - l) \, \pi }
                  {      n        }
    \\
    &=
    \sum_{ l = 0 }^{ 2 \, n - 1 }
      \mu_l \, n \left(
        \delta_{ p + l - 2 \, n, 0 }
        +
        \delta_{ p - l, 0 }
      \right)
    \\
    &=
    n \, \left( \mu_p + \mu_{ 2 \, n - p} \right)
    =
    2 \, n \, \mu_p.
  \end{aligned}
  $$
  This entails
  $$
  \begin{aligned}
    \mu_l
    &=
    \frac{    1   }
         { 2 \, n }
    \sum_{ k = 0 }^{ 2 \, n - 1 }
      \lambda_{ k + 1 } \,
      \cos \frac{ k \, l \, \pi }
                {      n        }
              \\
    &=
    \frac{    1   }
         { 2 \, n }
    \left[
      \lambda_1
      +
      (-1)^l \, \lambda_{n + 1}
      +
      2 \sum_{ k = 1 }^{ n - 1 }
        \lambda_{ k + 1 } \,
        \cos \frac{ k \, l \, \pi }
                  {      n        }
    \right],
  \end{aligned}
  $$
  where we have used Eq. \eqref{eq:lambda_reflection}
  in the last step.

  However, we are usually given only $\lambda_1, \dots, \lambda_n$
  without $\lambda_{n + 1}$.
  %
  Fortunately, we can deduce the latter from
  the condition that $\mu_n = 0$, which means
  $$
  \begin{aligned}
    0 = \mu_n
    =
    \frac{1}{n}
    \left[
      \frac{ \lambda_1 + (-1)^n \, \lambda_{n+1} }
           {               2                     }
      +
      \sum_{ k = 1 }^{ n - 1 }
      \lambda_{k+1} (-1)^k
    \right].
  \end{aligned}
  $$
  This allows us to solve for $\lambda_{ n + 1 }$,
  yielding Eq. \eqref{eq:lambdan}

  Some examples for checking.
  For $n = 2$,
  $\lambda_3
  = \mu_0 - 2 \, \mu_1
  = -(\lambda_1 - 2 \, \lambda_2)$,
  and
  $$
  \begin{aligned}
  \mu_0
  &=
  \lambda_2,
  \\
  \mu_1
  &=
  \frac{ \lambda_1 - \lambda_2 }
       {           2           }.
  \end{aligned}
  $$
  For $n = 3$,
  $\lambda_4
  = \mu_0 - 2 \, \mu_1 + 2 \, \mu_2
  = \lambda_1 - 2 \, \lambda_2 + 2 \, \lambda_3$,
  and
  $$
  \begin{aligned}
  \mu_0
  &=
  \frac{ 1 } { 3 }
  \left(
    \lambda_1 + 2 \, \lambda_3
  \right),
  \\
  \mu_1
  &=
  \frac{ 1 } { 3 }
  \left(
    \frac 3 2 \lambda_2
    -
    \frac 3 2 \lambda_3
  \right)
  \\
  \mu_2
  &=
  \frac{ 1 } { 3 }
  \left(
    \lambda_1
    -
    \frac 3 2 \lambda_2
    +
    \frac 1 2 \lambda_3
  \right)
  \end{aligned}
  $$
}
Thus, the algorithm of stabilization consists of three steps.
%
\begin{enumerate}
  \item
    Compute the eigenvalues from $\lambda_0, \dots, \lambda_{n-1}$
    from the given $\mu_0, \dots, \mu_b$ with $b < n$
    according to Eq. \eqref{eq:wband_eigenvalue}.
  \item
    Replace the negative eigenvalues by zeros.
  \item
    Compute the new $\mu_l$ $(0 \le l < n)$ from
    the modified eigenvalues according to
    Eq. \eqref{eq:mu_from_lambda}.
\end{enumerate}
%
Note that, after the stabilization,
the window is usually widened.
